---
title: Visualizing color frequency of clothing choices
author: Allen Chang
date: '2018-04-01'
slug: visualizing-color-frequency-of-clothing-choices
categories: []
tags:
  - clothing
  - visualization
  - R
output:
  blogdown::html_page:
    toc: true

---

In a preivous post, I visualizled color choices in clothing using slope charts and histograms. This time I wanted to get a bit more creative. In this post I'm going to explore how color frequencies in my clothing choices change across months by using waffle charts.

As usual, to start here's a list of the packages that I used in this post. 

```{r message=FALSE, warning=FALSE}
library(readr)
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)
```

### Data Preprocessing

The first step was to load in a csv of my clothing choices and the associated colors of each article of clothing. All I was interested this time was just the primary and secondary colors worn, so I selected those from my data and gathered them into one column. Before breaking up color counts by month, I wanted to see what the overall counts were over the accumulated September-March time periods.

```{r echo=TRUE, message=FALSE, warning=FALSE}
file_path <- "../../data/colors_3_2018.csv"

data <- read_csv(file_path)

color_tbl <- data %>%
  select(Color1, Color2) %>%
  gather(key, color) %>%
  filter(!color == "NA") %>%
  group_by(color) %>%
  summarise(color_count = n()) %>% 
  arrange(desc(color_count))

head(color_tbl, 10)
```

Woah, black is my most worn color with over time the number of wears than the next color white. As I move into thinking about frequencies of colors per month though, I wanted to make sure that I excluded colors that were only worn a few time. I created a dataframe of colors in each month that were worn less than 4 times and then anti-joined them with a dataframe of monthly color counts. 

```{r}
exclude_colors <- data %>%
  select(Color1, Color2, Day) %>%
  gather(key, color, -Day) %>%
  filter(!color == "NA") %>%
  mutate(Day = mdy(.$Day), month = floor_date(Day, unit = "month")) %>%
  group_by(month, color) %>%
  summarise(color_count = n()) %>%
  filter(color_count < 4)

head(exclude_colors)

color_tbl_month <- data %>%
  select(Color1, Color2, Day) %>%
  gather(key, color, -Day) %>%
  filter(!color == "NA") %>%
  mutate(Day = mdy(.$Day), month = floor_date(Day, unit = "month")) %>%
  select(month, color) 

color_tbl_final <- anti_join(color_tbl_month, exclude_colors, by = c("month", "color"))

head(color_tbl_final, 8)
```

### Constructing a grid of colors

Next I had to create the grid with which to plot my colors. I wrote a small function that calculated the frequencies of colors in each month and filled in a 10 x 10 grid. The resulting dataframe had columns corresponding to the month, the x and y coordinates, and the color for each box within the grid.

```{r}
color_freq_month <- plyr::ddply(color_tbl_final, .variables = c("month"), function(x){
  nrows <- 10
  df <- expand.grid(y = 1:nrows, x = 1:nrows)
  color_table <- data.frame(ceiling(table(x$color) * ((nrows*nrows)/(length(x$color)))))
  color <- rep(color_table$Var1, color_table$Freq)[1:100]
  cat_table <- table(color)
  df$color <- factor(rep(names(cat_table), cat_table))
  df
})

head(color_freq_month, 8)
```

### Visualizing color frequencies

Finally it was time to visualize my frequencies. I plotted my waffle chart using a geom_tile layer and manually assigned fill color using a vector of hexadecimal values I chose. 

```{r color_by_months, echo=TRUE, message=FALSE, warning=FALSE}
color_vec <- c("#000000", "#8C510A", "#08306B", "#FFFFFF", "#F6E8C3",
           "#4292C6", "#4D4D4D","#BABABA", "#006D2C", "#54278F", "#A50F15", "#F3EFE0", "#BF5120")

ggplot(color_freq_month, aes(x = x, y = y, fill = color)) +
  geom_tile(color = "gray18", size = 0.5) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), trans = 'reverse') +
  scale_fill_manual(name = "colors",
                    values = color_vec) +
  labs(title="Frequency distribution of colors worn") +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1), 
        strip.background = element_rect(color = "black", size = 1),
        plot.title = element_text(size = rel(1.2)),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_blank(),
        legend.position = "right") +
  facet_wrap(~month)
```

Waffle charts are a nice way of visualizing color because you can see they visualize proportions in a cleaner way than something like pie charts. If I was interested in seeing what percentage of my clothes worn in September of 2017 were black, all I would have to do is tally up the number of sqaares (4 x 10). Hopefully as it starts to warm in Boston though I might start trending back more towards lighter colors!

## References

[waffle charts](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#Waffle%20Chart)

[waffle package -not used here but a good reference-](https://github.com/hrbrmstr/waffle)

