---
title: "Network analysis of my daily clothing choices"
author: "Allen Chang"
date: '2018-01-06'
slug: daily_clothing_network
tags:
- R
- clothing
- networks
categories: []
---

This is a continuation of my previous analysis on tracking preferences in clothing choices (part 1 is [here](https://functionallydefined.netlify.com/posts/daily_wear/)). The inspiration for this post came about after re-reading David Robinson's network analysis of the movie [love actually](http://varianceexplained.org/r/love-actually-network/) and realizing that a similar approach could be applied to my sparse network of clothing choices. I found additional help by working through the Network Analysis course on [Datacamp](https://www.datacamp.com/courses/network-analysis-in-r).

As per usual, these were the packages used. 

```{r packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(reshape2)
library(igraph)
```

I used the same dataset as my previous analysis in which I recorded the maker and color for each article of clothing that I wore on each day. To fix possible errors in capitalization, I converted all character strings to upper case, and modified the day column to go from 1 to 93. I also filtered out rows from when it was too warm to wear a jacket.  

```{r read data, echo=TRUE, message=FALSE, warning=FALSE}
# Prepare data
wear_tbl <- read_csv("daily_wear.csv")

wear_tbl <- wear_tbl %>% 
  mutate(day = rep(c(1:93), each = 4), maker_up = toupper(Maker)) %>%
  select(day, maker_up) %>%
  filter(!maker_up == "NA")

wear_mat <- wear_tbl %>%
  acast(maker_up ~ day, fun.aggregate = length)
```

Following David Robinson, I elected to run a few visualizations to examine relationships between makers. The visualization was a  hierarchical cluster on my data using Canberra distance, which is a weighted form of the Manhattan distance. I opted to use normalized and non-normalized data to view how clustering differed. 

```{r normalize, echo=TRUE, message=FALSE, warning=FALSE}
# normalize data
norm <- wear_mat / rowSums(wear_mat)
```

```{r clust_plots, echo=FALSE, message=FALSE, warning=FALSE}
h <- hclust(dist(norm, method = "canberra"))
plot(h)

h <- hclust(dist(wear_mat, method = "canberra"))
plot(h)
```

For this data, I only recorded 3 months worth of choices, so some combinations of makers only occur once. I plan on continuing to gather data after my winter holiday, so it will be interesting to see how well the clustering works with more data. Again borrowing from David, I plotted the timeline of wears for each day.

```{r timeline_code, echo=FALSE, message=FALSE, warning=FALSE}
# Prepare data for timline vis
ordering <- h$labels[h$order]

day_vis <- wear_tbl %>%
  ungroup() %>%
  mutate(day = as.numeric(factor(day)),
         maker = factor(maker_up, levels = ordering))
```

```{r timeline_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(day_vis, aes(day, maker)) +
  geom_point() +
  geom_path(aes(group = day))
```

I really loved this timeline because it let me point out time related preferences. You can see that I heavily favored wearing Vans early in the fall, but as it cooled off I start to transition more into my other shoes. One drawback of this plot though is that it's a little diffiult to see strengths of connections between one maker and all other makers. Again following David, I plotted a heatmap of the adjacency matrix. 

```{r cooccurence, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
## cooccurence mat
cooccur <- wear_mat %*% t(wear_mat)

heatmap(cooccur, cexRow = 0.8, cexCol = 0.8)
```

That made tracking what connections each of my makers had and the strength of those connections much easier to see. Next, I used the igraph package as another way to plot the network. 

```{r igraph, echo=TRUE, message=FALSE, warning=FALSE}
g <- graph.adjacency(cooccur, weighted = TRUE, mode = "undirected", diag = FALSE)

# Places vertex labels around circular layout - code via stack, see references for post
radian.rescale <- function(x, start=0, direction=1) {
  c.rotate <- function(x) (x + start) %% (2 * pi) * direction
  c.rotate(scales::rescale(x, c(0, 2 * pi), range(x)))
}
lab.locs <- radian.rescale(x=1:25, direction=-1, start=0)
```

```{r network as circle, echo=FALSE, message=FALSE, warning=FALSE}
plot(g, 
     vertex.label.color = "black", 
     vertex.label.cex = .4,
     vertex.label.font = 2,
     vertex.label.dist= 0.75,
     vertex.label.degree=lab.locs,
     edge.color = 'gray77',
     vertex.size = 0,
     edge.arrow.size = 0.2,
     layout = layout.circle(g))

```

This was still a little too busy for me, so to clean things up I opted to remove vertices with less than 4 edges (if I only wore a maker once, it could have at max 3 other edges). 

```{r subset network, echo=TRUE, message=FALSE, warning=FALSE}
counts <- as.data.frame(lengths(as_adj_list(g))) %>%
  rownames_to_column()

names(counts) <- c("maker", "count")
counts <- counts %>%
  filter(count < 4)

g2 <- delete_vertices(g, counts$maker)
```

```{r subset network circle layout, echo=FALSE, message=FALSE, warning=FALSE}
plot(g2, 
     vertex.label.color = "black", 
     vertex.label.cex = .4,
     vertex.label.font = 2,
     vertex.label.dist= 0.5,
     edge.color = 'gray77',
     vertex.size = 0,
     edge.arrow.size = 0.2,
     layout = layout.circle(g2))
```

That looked better, but I still thought there might be a better layout that I could use. 

```{r fr / sphere vis, echo=FALSE, message=FALSE, warning=FALSE}
plot(g2,
     vertex.label.color = "black", 
     vertex.label.cex = .4,
     vertex.label.font = 2,
     vertex.label.dist= 0.5,
     edge.color = 'gray77',
     vertex.size = 0,
     edge.arrow.size = 0.2,
     layout = layout_with_fr(g2))

plot(g2,
     vertex.label.color = "black", 
     vertex.label.cex = .4,
     vertex.label.font = 2,
     vertex.label.dist= 0.5,
     edge.color = 'gray77',
     vertex.size = 0,
     edge.arrow.size = 0.2,
     layout = layout_on_sphere(g2))
```

These looked pretty good. I liked that the vertices with the most connections were placed inwards. Next, I moved on to look at some basic network attributes. First I wanted to look at how well connected all of the makers in my network were using edge density and the diameter of the graph.

```{r network attributes, echo=TRUE, message=FALSE, warning=FALSE}
# Get network density
gd <- edge_density(g)
gd

# Get network diameter
diameter(g, directed = FALSE)
```

An edge density of 0.35 is not surprising given the number of makers that were only worn once in my dataset. I further determined how connected the makers were by taking the average path length and comparing it to a random graph.

```{r message=FALSE, warning=FALSE}
# Get the average path length of my maker network
g.apl <- mean_distance(g, directed = FALSE)
g.apl

# Generate 1000 random graphs
gl <- vector('list', 1000)

for(i in 1:1000){
  gl[[i]] <- erdos.renyi.game(n = gorder(g), p.or.m = gd, type = "gnp")
}

# Calculate average path length of 1000 random graphs
gl.apl <- lapply(gl, mean_distance, directed = FALSE)
gl.apls <- unlist(gl.apl)

# Calculate the proportion of random graphs with an average path length lower than my maker network
sum(gl.apl < gl.apls)/1000
```

```{r path length comparison, echo=FALSE, message=FALSE, warning=FALSE}
# Plot the distribution of average path lengths
hist(gl.apls, xlim = range(c(1.5, 2)))
abline(v = mean_distance(g, directed = FALSE), col = "red", lty = 3, lwd=2)
```

To see how well adjacent nodes were interconnected in subgroups, I looked at the transitivity and again compared the probability against a null graph. 

```{r transitivity, message=FALSE, warning=FALSE}
# Calculate the global transitivity of the network.
transitivity(g)

# Calculate average transitivity of 1000 random graphs
gl.tr <- lapply(gl, transitivity)
gl.trs <- unlist(gl.tr)

# Get summary statistics of transitivity scores
summary(gl.trs)

# Proportion of random graphs with a transitivity score higher than my maker network.
sum(gl.trs > gl.tr)/1000
```

It's pretty clear from my network measures that my network is more interconnected than the random networks. That means I might not just be choosing clothes at random everyday! Maybe some makers do pair well with each other. Finally, I ran some community detection to understand the structure of my network. I used two types of community detection functions, fast-greedy communities and edge-betweenness communities.  

```{r community detection, echo=FALSE, message=FALSE, warning=FALSE}
# Run fast-greedy community detection
kc = fastgreedy.community(g)

# Get sizes of each community
sizes(kc)

# Run edge-betweenness community detection
gc = edge.betweenness.community(g)

# Get sizes of each community
sizes(gc)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot community networks side-by-side
par(mfrow = c(1, 2)) 
plot(kc, g,
     vertex.label = NA)
plot(gc, g, 
     vertex.label = NA)
```

Looking forward, I still have one more post planned for this dataset using the [threejs](https://cran.rstudio.com/web/packages/threejs/index.html) package to create interactive networks. I've had a lot of fun working through this data, and I'm looking forward to continuing to track and think about my clothing choices.

### References
[labels around circular layout](https://stackoverflow.com/questions/23209802/placing-vertex-label-outside-a-circular-layout-in-igraph/23210025)

[david robinson's love actually network](http://varianceexplained.org/r/love-actually-network/)

[datacamp - network analysis in R](https://www.datacamp.com/courses/network-analysis-in-r)
